---
  output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# GoNoGo

<!-- badges: start -->

<!-- badges: end -->

The goal of `GoNoGo` is to provide the posterior probability and operating characteristics of Bayesian Go/No-Go decision-making based on specific model assumptions. Currently, it specifically covers scenarios involving single normal mean, two independent normal mean differences, and two independent normal mean ratio analysis, considering both informative and non-informative priors. For a more in-depth understanding of the technical aspects, please refer to the `Vignette`.

## Dependencies

`GoNoGo` requires the following `R` package: `devtools` (for installation only). Please install it before installing `GoNoGo`, which can be done as follows (execute from within a fresh `R` session):

``` r
install.packages("devtools")
library(devtools)
```


## Installation

Once the dependencies are installed, `GonoGo` can be loaded using the following command:

``` r
devtools::install_github("Merck/GoNoGo")
library(GoNoGo)
```

## Scenarios (In current version)

* Operating characteristics
  + Normal 1 sample - Prior: non-informative - Hypothesis: greater than, less than, between
  + Normal 1 sample - Prior: informative - Hypothesis: greater than, less than, between
  + Normal 2 sample - Difference/ratio - Prior: non-informative - Hypothesis: greater than, less than, between
  + Normal 2 sample - Difference/ratio - Prior: informative - Hypothesis: greater than, less than, between
* Posterior probability
  + Normal 1 sample - Prior: non-informative - Hypothesis: greater than, less than, between
  + Normal 1 sample - Prior: informative - Hypothesis: greater than, less than, between
  + Normal 2 sample - Difference/ratio - Prior: non-informative - Hypothesis: greater than, less than, between
  + Normal 2 sample - Difference/ratio - Prior: informative - Hypothesis: greater than, less than, between
* Plot

## Example 1 (Operating characterstics)

This is a basic example which shows you how to generate operating characteristics for a normal one sample non-informative prior with hypothesis = "greater than":

```{r example1}
##################
# Load libraries #
##################

library(GoNoGo)
library(ggplot2)
library(grid)
library(gridExtra)

n.t = rep(8, 9) # Sample size in treatment arm
mu.t = rep(c(40, 45, 50), each = 3) # True mean of the treatment arm
sd.t = rep(15, 9) # True sd of the treatment arm
hypothesis = "between" # "greater than", "less than", "between"
threshold =  matrix(rep(c(35, 45), each = 9), ncol = 2) # k: Target region cut-off 
PP.cutoffGo = rep(c(0.6, 0.7, 0.8), 3) # c1: Cut-off for probability of Go
PP.cutoffNoGo = rep(c(0.6, 0.7, 0.8), 3) # c2: Cut-off for probability of No-Go
prior = "non-informative" # Prior choice
reps = 1000 # Number of repetitions for samples from the posterior using simulation
nsim = 1000 # Number of simulations

set.seed(1234)
pos.val = vector("list", length(n.t))
for(i in 1:length(n.t)){
  pos.val[[i]] = Normal1SampleGNG(n.t = n.t[i], mu.t = mu.t[i], sd.t = sd.t[i], 
                                  hypothesis = hypothesis, threshold = threshold[i,],
                                  PP.cutoffGo = PP.cutoffGo[i], 
                                  PP.cutoffNoGo = PP.cutoffNoGo[i],
                                  nsim = nsim, reps = reps)$pos
}
data = dataoutput(prob.val = pos.val)
data$mu.t = unlist(data$mu.t)
data$PP.cutoffGo = unlist(data$ PP.cutoffGo)

data$mu.t = as.factor(data$mu.t)
data$PP.cutoffGo = as.factor(data$ PP.cutoffGo)

data
```

```{r example1.plot}
p1 = GNGplot(data = data, x.var = data$mu.t,
                     fill.var = data$PP.cutoffGo,
                     hypothesis = data$hypothesis,
                     x.lab = "mu.t",
                     title.text =  paste("PP(True mean is", hypothesis, 
                                         "threshold) >= PP cutoff"),
                     legend.title.text = "PP.cutoffGo")

p2 = GNGplot(data = data, x.var = data$PP.cutoffGo,
                     fill.var = data$mu.t,
                     hypothesis = data$hypothesis,
                     x.lab = "PP.cutoffGo",
                     title.text =  paste("PP(True mean is", hypothesis, 
                                         "threshold) >= PP cutoff"),
                     legend.title.text = "mu.t" )

grid.arrange(p1, p2, nrow = 1)
```


## Example 2 (Posterior probability)

This is a basic example which shows you how to generate posterior probability based on the previous parameters.

```{r example2}
set.seed(1234)
pp.val = vector("list", length(n.t))
for(i in 1:length(n.t)){
  pp.val[[i]] = Normal1SamplePP(n.t = n.t[i], mu.t = mu.t[i], sd.t = sd.t[i],
                                hypothesis = hypothesis, 
                                threshold = threshold[i,], prior = prior, 
                                PP.cutoffGo = PP.cutoffGo[i], 
                                PP.cutoffNoGo = PP.cutoffNoGo[i], reps = reps)
}
posterior.prob = dataoutput(prob.val = pp.val, output.format = "PP")
posterior.prob
```

# Work in progress
Currently, the followings are in progress

* R-shiny app corresponding to Normal
* Binary Go/No-Go.
